---
title: Quantifying reproducibility - Results from a scoping review on reproducibility metrics and simulation studies into their real-world applicability <br>
subtitle: Conference of the International Society for Clinical Biostatistics - August 26, 2025
author: Rachel Heyard*, Samuel Pawel, Joris Frese, Bernhard Voelkl, Hanno Würbel, Sarah McCann, Leonhard Held, Kimberley Wever, Helena Hartmann, Luise Townsin, and Stephanie Zellers
institute: Center for Reproducible Science
preview-links: true
format:
  beamer:
    theme: uzh
    aspectratio: 169
    include-in-header: beamerthemeuzh.sty
    bibliography: refs.bib
    link-citations: true
    urlcolor: purple
    inclemental: true
---

```{r setup}
library(knitr)

# Load libraries
library(kableExtra)
library(here)
# library(wesanderson)
library(gridExtra)
library(tidyverse)
library(ERforResearch)
library(stringr)
library(readr)

# Source needed functions

opts_chunk$set(echo = FALSE, out.extra = "", warning = FALSE,
               message = FALSE, cache.lazy = FALSE)
options(knitr.kable.NA = "---")

re_sample <- FALSE
```


## Part of improving Reproducibility In SciencE (iRISE)



```{=latex}
\begin{minipage}{0.7\textwidth}
{\footnotesize
iRISE receives funding from the European Union's Horizon Europe research and innovation programme under grant agreement No 101094853. Views and opinions expressed are however those of the author(s) only and do not necessarily reflect those of the European Union or the European Research Executive Agency (ERA). Neither the European Union nor the ERA can be held responsible for them.}
\end{minipage}%
\hfill
\begin{minipage}{0.2\textwidth}
  \centering
  \includegraphics[width=\linewidth]{img/iRISE_logo_dark_round.png}
\end{minipage}
```

<br>
<br>
<br>

```{=latex}
\begin{center}
\begin{minipage}{0.3\textwidth}
  \centering
  \includegraphics[width=\linewidth]{img/EN_Co-fundedbytheEU_RGB_POS.png}
\end{minipage}
\hspace{0.01\textwidth}
\begin{minipage}{0.3\textwidth}
  \centering
  \includegraphics[width=\linewidth]{img/UKRI-logo.png}
\end{minipage}
\hspace{0.01\textwidth}
\begin{minipage}{0.3\textwidth}
  \centering
  \includegraphics[width=\linewidth]{img/WBF_SBFI_EU_Frameworkprogramme_E_RGB_pos_quer.jpg}
\end{minipage}
\end{center}
```


## What do we mean by "Reproducibility"?

**Definition of reproducibility, for the purpose of this presentation**:

<br>

*Reproducibility* = results of studies are consistent/agree.
  
  <br><br>
  
  \pause
  
- *Computational reproducibility*:  
  Using the *same data and code* to reproduce the original results.

- *Results reproducibility*:  
  New data is collected and analysed using similar methods and materials.

- *Partial reproducibility*:  
  Only parts of the study are reproduced (e.g., experiment, data collection, analysis), and results are compared.
  
  
## The million-dollar question


```{=latex}
\begin{itemize}\setlength{\itemsep}{1.5em}
  \item \emph{When do results agree?}
  \item \emph{What does it mean for study results to be consistent?}
\end{itemize}
```


## Reproducibility Metrics Scoping Review | Question

**Research questions:**

```{=latex}
\begin{itemize}
\setlength{\itemsep}{1.5em}
  \item \only<1>{Which metrics have been used or suggested to quantify or assess which type of reproducibility?}
        \only<2->{\textbf{Which metrics have been used or suggested to quantify or assess which type of reproducibility?}}
  \item \only<1>{Which ones were suggested only, or have a clear implementation?}
        \only<2->{\textcolor{gray}{Which ones were suggested only, or have a clear implementation?}}
  \item \only<1>{How can they be interpreted and in which situations are they useful?}
        \only<2->{\textcolor{gray}{How can they be interpreted and in which situations are they useful?}}
  \item \only<1>{What are assumptions and limitations?}
        \only<2->{\textcolor{gray}{What are assumptions and limitations?}}
\end{itemize}
```

## Reproducibility Metrics Scoping Review | Methods

**Scoping Review in Two Parts**

* \underline{Application papers} -- For reproducibility metrics **used in practice**, a list of large-scale reproducibility projects was compiled by the project team.
  
  \pause
  ::: {.callout-note} 
  # Large-scale reproducibility projects
  Large-scale efforts to measure the reproducibility of a field, method or type of study.
  :::
  
  <!-- add blank line here -->
  \pause
  
* \underline{Methodological papers} -- For reproducibility metrics **proposed, suggested or discussed**, a systematic search was conducted to identify relevant literature.


## Application papers

- **49 records** of reproducibility projects across fields of research (based on OpenAlex):
  - Health and life sciences -- 10 (20.4%)
  - Physical sciences -- 6 (12.2%)
  - Social sciences -- 33 (67.3%)

<!-- add blank line here -->
<!-- add blank line here -->

- Categorised into three project types: 
  1. **Many phenomena, one study**: many original hypotheses tested, each hypothesis once;
  2. **One phenomenon, many studies**:one hypothesis is tested by many different teams;
  3. **Many phenomena, many studies**: many hypotheses are tested, each multiple times


## Metrics used in application papers
- Various metrics were used to assess reproducibility:
  - Some compared **statistical significance**
  - Others compared **effect sizes**
  - Some used **meta-analyses**
  - Others relied on **subjective assessments**
  - *some used others*



<!-- **Frequency of metric types in application papers:** -->

| Type of Metric                      | N (%)      |
|------------------------------------|------------|
| Agreement in statistical significance | 32 (65.3%) |
| Agreement in effect size              | 35 (71.4%) |
| Meta-analysis of study results       |  9 (18.4%) |
| Subjective assessment                | 14 (28.6%) |


## Number of metrics used in application papers

```{r fig.align='center', out.height='90%'}
# Replace with your actual plot path or code
knitr::include_graphics("img/count_metrics.pdf")
```

