---
title: "Let's talk about uncertainty in funding allocation"
subtitle: "ESO 2023 Workshop - Peer Review Under Review"
author: "_Rachel Heyard_"
date: 'Wednesday, February 8th 2022'
output: 
  ioslides_presentation: 
    incremental: true
    widescreen: true
    css: styles.css
    logo: UZH_CRS.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "figures/",
                      warning = FALSE, message = FALSE) 

library(ERforResearch)
library(kableExtra)
library(stringr)
library(readr)

re_sample <- FALSE 
```


```{r get-data}
# Load the data from Zenodo:
path_to_xlsx <- "https://zenodo.org/record/4531160/files/individual_votes.xlsx"
stem_mat <- openxlsx::read.xlsx(xlsxFile = path_to_xlsx, sheet = "pm_stem")
stem <- stem_mat %>% 
  get_right_data_format(prefix_assessor = "voter")

assessor_coi <-
  stem %>% 
  group_by(assessor) %>% 
  count() %>% 
  mutate(assessor_coi = n != 18) %>% 
  select(-n)
proposal_coi <-
  stem %>% 
  group_by(proposal) %>% 
  count() %>% 
  mutate(proposal_coi = n != 29) %>% 
  select(-n)

stem <- stem %>% 
  left_join(assessor_coi) %>% 
  left_join(proposal_coi) 
```

```{r some-helper-functions}
## Some helper functions for later:
# Function to extract the Intraclass correlation coefficient (ICC) from the 
# matrix of individual votes:
get_icc_from_matrix <- function(individual_votes) {
  (individual_votes %>% 
     select(-proposal) %>% 
     mutate_all(function(x) ERforResearch:::get_num_grade_snsf(x)) %>% 
     as.data.frame() %>% 
     psych::ICC(x = ., missing = FALSE))$results %>% 
    filter(type == "ICC3") %>% # A fixed set of k judges rates each target
    mutate(icc = paste0(round(ICC, 2), " (",
                        round(`lower bound`, 2), "; ", 
                        round(`upper bound`, 2), ")")) %>% 
    pull(icc) %>% 
    return()
}

# Function to graphically represent all the votes, and the ICC from the
# matrix of individual votes:
get_grades_plot <- function(long_individual_votes, individual_votes_mat = NULL,
                            x_min = NULL, x_max = NULL, title = "",
                            jitter_h = .02, jitter_w = .025,
                            jitter_alpha = .5){
  if (is.null(x_min)){
    x_min <- long_individual_votes %>% pull(num_grade) %>% min
    x_min <- ifelse(x_min > 1, x_min - 1.2, x_min -.2)
  }
  if (is.null(x_max)){
    x_max <- long_individual_votes %>% pull(num_grade) %>% max
    x_max <- ifelse(x_max < 6, x_max + 1.2, x_max +.2)
  }
  if (!is.null(individual_votes_mat)){
    icc <- get_icc_from_matrix(individual_votes_mat)
    title <- paste0(ifelse(title == "", "ICC = ",
                           paste0(title, ": ICC = ")), icc)
  }
  plot_data <- long_individual_votes %>% 
    group_by(proposal) %>% 
    mutate(avg = mean(num_grade, na.rm = TRUE)) %>% 
    ungroup() 
  plot_data <- plot_data %>% 
    select(proposal, avg) %>% 
    distinct() %>% 
    arrange(avg) %>% 
    mutate(order = 1:n()) %>% 
    select(-avg) %>% 
    left_join(plot_data, by = "proposal")
  plot_data %>% 
    ggplot(aes(x = num_grade, y = order)) + 
    geom_jitter(aes(color = "Individual Votes"), height = jitter_h, 
                width = jitter_w, alpha = jitter_alpha, size = 1) +
    geom_point(aes(x = avg, color = "Average"), size = 1.5) + 
    theme_minimal() +
    scale_y_continuous(breaks = unique(plot_data$order),
                       labels = 
                         as.numeric(str_extract_all(unique(plot_data$proposal),
                                                    "[0-9]+"))) +
    labs(x = "Assessor Vote", y = "Ordered proposal", title = title) +
    lims(x = c(x_min, x_max)) +
    scale_color_manual(name = " ", 
                       values = c("Average"="#FF6F6F",
                                  "Individual Votes" = "#848383")) +
    theme(axis.text.y = element_blank(),
          legend.position = c(0.5, .5),
          axis.title.y = element_blank(),
          panel.grid.major.y = element_line(size = .1))
}


```


## FYI

These slides only include the Figures and data analysis for the keynote I give/gave at the ESO 2023 Workshop on "Peer Review Under Review".

The complete slides are [here](https://osf.io/3r7et).




## The STEM panel of fellowship - funding instrument

```{r}

nb_cois <- sum(apply(stem_mat, 1, function(x) sum(is.na(x))))

stem %>% 
  summarise("Total nb of votes" = n(),
            "Nb proposals discussed" = n_distinct(proposal),
            "Nb panel members" = n_distinct(assessor)) %>% 
  mutate("Nb fundable proposals" = 6,
         "Nb COIs" = nb_cois,  
         "Scale used" = "A, AB, B, BC, C, D") %>% 
  kable(booktab = TRUE) %>%
  kable_classic_2()


```


## The STEM panel of fellowship - funding instrument
```{r barplot, fig.width=7}
plot_data <- stem %>% 
    group_by(proposal) %>% 
    mutate(avg = mean(num_grade, na.rm = TRUE)) %>% 
    ungroup() %>% 
    select(proposal, avg) %>% 
    distinct() %>% 
    arrange(avg) %>% 
    mutate(order = 1:n()) %>% 
    # select(-avg) %>% 
    left_join(stem, by = "proposal")

plot_data %>% 
  group_by(grade) %>% 
  summarise(count = n() ) %>% 
  ggplot(aes(x = grade, y = count)) +
  geom_col(fill = "#18ac7a", color = "white") +
  labs(y = "Count", x = " ") + 
  theme_classic()

```


## Variation within proposals 1


```{r variation-by-proposals, fig.width=10}
set.seed(1)
plot_data %>% 
  ggplot(aes(x = as.factor(order), y = num_grade)) +
  geom_violin(fill = alpha("#18ac7a", .6), linewidth = 1, 
              color = alpha("#18ac7a", .6)) +
  geom_jitter(size = 1.5, height = .15, color = "#A3A3A3",
              width = .2, alpha = .6) + 
  # geom_point(data = plot_data %>% select(order, avg) %>% distinct(),
             # aes(y = avg), col = "#9A1542", size = 6, pch = 13) +
  theme_classic() + 
  # scale_color_manual(values = c("#A3A3A3", "black")) +
  labs(y = "Assessor Vote", x = "Ordered proposal") +
  scale_y_continuous(breaks = 1:6,
                     labels = c("D", "C", "BC", "B", "AB", "A"),
                     limits = c(1, 6.2)) +
  theme(axis.text.x = element_blank(),
        legend.position = "none")
```


## Variation within proposals 2


```{r variation-by-proposals2, fig.width=10}
set.seed(1)
plot_data %>% 
  ggplot(aes(x = as.factor(order), y = num_grade)) +
  geom_violin(fill = alpha("#18ac7a", .6), linewidth = 1, 
              color = alpha("#18ac7a", .6)) +
  geom_jitter(size = 1.5, height = .15, aes(color = proposal_coi), #color = "#A3A3A3",
              width = .2, alpha = .6) + 
  # geom_point(data = plot_data %>% select(order, avg) %>% distinct(),
             # aes(y = avg), col = "#9A1542", size = 6, pch = 13) +
  theme_classic() + 
  scale_color_manual(values = c("#A3A3A3", "black")) +
  labs(y = "Assessor Vote", x = "Ordered proposal") +
  scale_y_continuous(breaks = 1:6,
                     labels = c("D", "C", "BC", "B", "AB", "A"),
                     limits = c(1, 6.2)) +
  theme(axis.text.x = element_blank(),
        legend.position = "none")
```

## Variation within proposals 3


```{r variation-by-proposals3, fig.width=10}
set.seed(1)
plot_data %>% 
  ggplot(aes(x = as.factor(order), y = num_grade)) +
  geom_violin(fill = alpha("#18ac7a", .6), linewidth = 1, 
              color = alpha("#18ac7a", .6)) +
  geom_jitter(size = 1.5, height = .15, aes(color = proposal_coi), #color = "#A3A3A3",
              width = .2, alpha = .6) + 
  geom_point(data = plot_data %>% select(order, avg) %>% distinct(),
             aes(y = avg), col = "#9A1542", size = 6, pch = 13) +
  theme_classic() + 
  scale_color_manual(values = c("#A3A3A3", "black")) +
  labs(y = "Assessor Vote", x = "Ordered proposal") +
  scale_y_continuous(breaks = 1:6,
                     labels = c("D", "C", "BC", "B", "AB", "A"),
                     limits = c(1, 6.2)) +
  theme(axis.text.x = element_blank(),
        legend.position = "none")
```

## Variation within assessors


```{r variation-by-assessor, fig.width=10}
set.seed(1)
plot_data %>% 
  ggplot(aes(x = assessor, y = num_grade)) +
  geom_violin(fill = alpha("#18ac7a", .6), linewidth = 1, 
              color = alpha("#18ac7a", .6)) +
  geom_jitter(size = 1.5, height = .15, color = "#A3A3A3",
              width = .2, alpha = .6) + 
  theme_classic() + 
  labs(y = "Assessor Vote", x = "Assessors") +
  scale_y_continuous(breaks = 1:6, labels = c("D", "C", "BC", "B", "AB", "A"),
                     limits = c(1, 6.2)) +
  theme(axis.text.x = element_blank(),
        legend.position = "none")
```

## Variation within assessors 2

```{r variation-by-assessor2, fig.width=10}
set.seed(1)
plot_data %>% 
  ggplot(aes(x = assessor, y = num_grade)) +
  geom_violin(fill = alpha("#18ac7a", .6), linewidth = 1, 
              color = alpha("#18ac7a", .6)) +
  geom_jitter(size = 1.5, height = .15, aes(color = assessor_coi),
              width = .2, alpha = .6) + 
  theme_classic() + 
  labs(y = "Assessor Vote", x = "Assessors") +
  scale_color_manual(values = c("#A3A3A3", "black")) +
  scale_y_continuous(breaks = 1:6, labels = c("D", "C", "BC", "B", "AB", "A"),
                     limits = c(1, 6.2)) +
  theme(axis.text.x = element_blank(),
        legend.position = "none")
```


```{r useERforResearch-get-mcmc}
if (re_sample){
  stem_mcmc <- get_mcmc_samples(data = stem, 
                                id_proposal = "proposal",
                                id_assessor = "assessor", 
                                grade_variable = "num_grade", 
                                n_iter = 100000,
                                n_burnin = 8000, 
                                n_adapt = 2000,
                                names_variables_to_sample = 
                                  c("proposal_intercept", "tau_proposal", 
                                    "tau_assessor", "rank_theta", 
                                    "nu"))
  write_rds(stem_mcmc, file = "stem_mcmc.rds")
}

stem_mcmc <- readRDS("stem_mcmc.rds")
```


```{r plot-results}


plot_assessor_stem <- 
  assessor_behavior_distribution(get_mcmc_samples_result = stem_mcmc,
                                 names_assessors = "assessor",
                                 n_assessors = stem %>% 
                                   summarise(n_distinct(assessor)) %>% pull(),
                                 xlim_min = -1.25,
                                 xlim_max = 1.05,
                                 scale = 2.5)

plot_proposal_stem <-
  assessor_behavior_distribution(get_mcmc_samples_result = stem_mcmc,
                                 names_assessors = "proposal",
                                 name_mean = "proposal_intercept",
                                 n_assessors = stem %>%
                                   summarise(n_distinct(proposal)) %>% pull(),
                                 xlim_min = -1.01, xlim_max = .8,
                                 scale = 2.5)

ranks_stem <- get_er_from_jags(data = stem,
                               id_proposal = "proposal",
                               id_assessor = "assessor", 
                               grade_variable = "num_grade",
                               mcmc_samples = stem_mcmc)

plot_stem <-  
  plotting_er_results(ranks_stem, result_show = TRUE, title = "STEM",
                      id_proposal = "id_proposal",
                      pt_size = .5, line_size = .2, draw_funding_line = FALSE, 
                      line_type_fl = "longdash", 
                      color_fl = "darkgray", grep_size = 3,
                      how_many_fundable = 6) 

plot_er_dist_stem <- 
  plot_er_distributions(get_mcmc_samples_result = stem_mcmc,
                        n_proposals = stem %>% 
                          summarise(n_distinct(proposal)) %>% pull(),
                        name_er = "rank_theta", title = " ",
                        number_fundable = 6,
                        outer_show = FALSE, proposal = "")

```

## Modeling the assessor behaviour
```{r variation-by-assessor-modeled, fig.width=5}
plot_assessor_stem
```



## Modeling the proposal quality
```{r variation-by-proposals-modeled, fig.width=5}
plot_proposal_stem
```


## Modeling the ranking
```{r modeled-ranks, fig.width=10}
plot_stem
```

## Recommendation
```{r br-recommendation, fig.width=10}
plot_er_dist_stem
```


## Rankogram
```{r}

plot_rankogram <- function(data, cumulative_rank_prob = FALSE, id_proposal, id_assessor, 
                           mcmc_samples = NULL, grade_variable = "num_grade", path_to_jags_model = NULL, 
                           n_chains = 4, n_iter = 10000, n_burnin = 4000, n_adapt = 1000, 
                           id_panel = NULL, max_iter = 1e+06, theta_name = "proposal_intercept", 
                           tau_name_proposal = "tau_proposal", assessor_behavior_name = "nu", 
                           sigma_name = "sigma", tau_name_assessor = "tau_assessor", 
                           tau_name_panel = "tau_panel", rank_theta_name = "rank_theta", 
                           assessor_name = "assessor_intercept", ordinal_scale = FALSE, 
                           point_scale = NULL, heterogeneous_residuals = FALSE, seed = 1991, 
                           quiet = FALSE, dont_bind = FALSE, inits_type = "random", 
                           names_variables_to_sample = NULL, initial_values = NULL, 
                           rhat_threshold = 1.01, runjags_method = "parallel") 
{
  n_proposal <- data %>% dplyr::pull(get(id_proposal)) %>% 
    unique() %>% length()
  overall_mean <- data %>% mutate(num_grade = get(grade_variable)) %>% 
    group_by(get(id_proposal)) %>% summarise(av = mean(.data$num_grade, 
                                                       na.rm = TRUE)) %>%
    dplyr::pull(.data$av) %>% mean()
  if (is.null(mcmc_samples)) {
    mcmc_samples <- 
      get_mcmc_samples(data = data, id_proposal = id_proposal, 
                       id_assessor = id_assessor, grade_variable = grade_variable, 
                       path_to_jags_model = path_to_jags_model, n_chains = n_chains, 
                       n_iter = n_iter, n_adapt = n_adapt, n_burnin = n_burnin, 
                       max_iter = max_iter, id_panel = id_panel, theta_name = theta_name, 
                       rank_theta_name = rank_theta_name, assessor_name = assessor_name, 
                       assessor_behavior_name = assessor_behavior_name, 
                       tau_name_proposal = tau_name_proposal, tau_name_assessor = tau_name_assessor, 
                       tau_name_panel = tau_name_panel, sigma_name = sigma_name, 
                       ordinal_scale = ordinal_scale, point_scale = point_scale, 
                       heterogeneous_residuals = heterogeneous_residuals, 
                       inits_type = inits_type, initial_values = initial_values, 
                       names_variables_to_sample = names_variables_to_sample, 
                       seed = seed, quiet = quiet, rhat_threshold = rhat_threshold, 
                       runjags_method = runjags_method)
  }
  else {
    if (length(mcmc_samples) != 7) {
      stop(paste0("Make sure that the object given to mcmc_samples is an ", 
                  "object that was build with get_mcmc_samples()."))
    }
  }
  colnames_rank_theta <- paste0(rank_theta_name, "[", seq_len(n_proposal), 
                                "]")
  if (is.list(mcmc_samples$samples)) {
    mcmc_samples_rank_thetas <- 
      do.call(rbind, mcmc_samples$samples$mcmc)[, colnames_rank_theta]
  }
  else {
    mcmc_samples_rank_thetas <- mcmc_samples$samples[, colnames_rank_theta]
  }
  p_j_b <- matrix(NA, nrow = n_proposal, ncol = n_proposal)
  for (i in seq_len(n_proposal)) {
    for (j in seq_len(n_proposal)) {
      p_j_b[i, j] <- mean(mcmc_samples_rank_thetas[, i] == 
                            j)
    }
  }
  sucra <- sapply(seq_len(nrow(p_j_b)), function(i) {
    mean(cumsum(p_j_b[i, -nrow(p_j_b)]))
  })
  colnames(p_j_b) <- seq_len(ncol(p_j_b))
  rownames(p_j_b) <- data %>% dplyr::pull(get(id_proposal)) %>% 
    unique()
  if (!cumulative_rank_prob) {
    p_j_b %>% as_tibble(rownames = "application") %>% 
      gather("rank", "prob", 
             paste0(1):paste0(ncol(p_j_b))) %>%
      mutate(rank = as.numeric(.data$rank)) %>% 
      ggplot(aes(x = .data$rank, y = .data$prob)) +
      geom_line() + 
      labs(x = "Rank", y = "Probability") +
      theme_minimal() + 
      facet_wrap(~.data$application)
  }
  else {
    sucra <- sapply(seq_len(nrow(p_j_b)), function(i) {
      mean(cumsum(p_j_b[i, -nrow(p_j_b)]))
    })
    names(sucra) <- data %>% dplyr::pull(get(id_proposal)) %>% 
      unique()
    p_j_b %>% as_tibble(rownames = "application") %>%
      gather("rank", "prob", paste0(1):paste0(ncol(p_j_b))) %>%
      mutate(rank = as.numeric(.data$rank)) %>% 
      group_by(.data$application) %>% 
      arrange(rank) %>% 
      # mutate(cum_prob = cumsum(.data$prob), sucra = sucra[head(.data$application, 
      #                                                          1)]) %>%
      ungroup() %>% ggplot(aes(x = .data$rank, 
                               y = .data$cum_prob)) + 
      geom_line() +
      labs(x = "Rank", y = "Cumulative probability") + 
      # geom_text(aes(x = n_proposal - 3, y = 0.5,
      #               label = paste0("SUCRA: ", head(round(sucra, 2), 1))), 
      #           size = 2) + 
      theme_minimal() +
      facet_wrap(~.data$application) 
  }
}

rankogram <- 
  plot_rankogram(data = stem,
                 id_proposal = "proposal",
                 id_assessor = "assessor", 
                 grade_variable = "num_grade",
                 rank_theta_name = "rank_theta",
                 cumulative_rank_prob = FALSE,
                 mcmc_samples = stem_mcmc) +
  theme(plot.title = element_blank())


rankogram
```

